### 2013-12-12 paper/paper.tex
rewrote assessment and conclusion

\documentclass[twocolumn]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{fancyvrb}

\title{Modularizing Street Map Generation \\
    \vspace{8pt} \large An Adventure in Tensors}
\author{William Clarkson \and Marcella Hastings \and Nathaniel Tenczar}
\date{\today}

\newcommand{\sqmat}[4]{\ensuremath{
    \left(\begin{array}{cc}
        #1 & #2 \\
        #3 & #4
    \end{array}\right)}}
\newcommand{\mkvec}[2]{\ensuremath{
    \left(\begin{array}{c}
        #1 \\
        #2 \\
    \end{array}\right)}}
\newcommand{\pt}{\textbf{p}}
\newcommand{\todo}[1]{\begin{center}\fbox{\parbox{150pt}{#1}}\end{center}}
\def \cpp {C\nolinebreak[4]\hspace{-.05em}\raisebox{.4ex}{\tiny\bf ++}~}

\begin{document}
\maketitle

\begin{abstract}
This paper addresses the issue of creating modular, extensible code given an
existing solution to a problem, namely, procedurally modeling large street
networks. We describe a functional implementation of a framework that allows
users to create a realistic street network over an empty domain, based on the
techniques described in \textit{Interactive Procedural Street Modelling}
\cite{chen}. In particular, we were able to express the underlying mathematics
more intuitively and clearly than is possible in an imperative language.
\end{abstract}

\section{Introduction}
Visual artists often face the time-consuming task of generating extensive,
realistic scenes by hand for movies and video games. However, the apparent
randomness found in systems like street layouts can be modelled effectively
with a variety of mathematical algorithms, giving rise to the growing field of
procedural generation. In this paper we will discuss an algorithm using tensor
fields, from \cite{chen}, to generate realistic city street maps. To facilitate
the generation of aesthetically pleasing maps which fit an artist's
requirements, we implement a simple system by which the user can input
constraints to the algorithm, shaping the resulting street map to fit their
needs.

\section{Existing Work}
\textit{Interactive Procedural Street Modelling} \cite{chen} introduces a novel
new concept of using tensor fields to model realistic street layouts. They
offer a well-defined pipeline of mathematical operations to go from simple user
inputs to a final street graph. They also implement a number of operations to
modify the street graph more granularly once it has been generated, such as the
removal and retracing of the roads in a defined region with a new
tensor field. The mathematical concepts fundamental to the procedural
generation algorithm are covered thoroughly and clearly.

While all of the information presented in the paper completely describes all of
the necessary mathematics for the algorithm in a clear manner, we found the
provided implementation more difficult to grasp. The large quantity of \cpp
code necessary to implement the full-featured interactive system provided by
Chen et al. made it difficult for us to thoroughly understand the constituent
components such that we would be able to modify or extend the existing code.

\section{Our Goals}
We approached the problem with the idea that we could use the power and
expressiveness of a functional language to produce a more modular, extensible,
and readable implementation of the core of the algorithm described in Chen et
al. which would provide a more accessible starting point for later work to
build off of. While we sacrifice the performance offered by the \cpp
implementation, which is closely tied to OpenGL, we contend that the gains in
programmer productivity will alleviate this deficit. While a production
implementation of such a pipeline should certainly be implemented with
efficiency in mind, we believe that, in the process of investigating new
algorithms, the ease with which the implementor can translate an idea or
mathematical concept into proof-of-concept code is of the utmost importance.
Later in the paper, we will assess our success in achieving these goals.

\section{Mathematical Background}\label{sec:math}
In this section we will introduce the mathematical concepts necessary to
understand our algorithm. A tensor $t$ is a geometric object represented by a
$2\times2$ symmetric, traceless matrix:
\[
    \sqmat{\cos{2\theta}}{\sin{2\theta}}{\sin{2\theta}}{-\cos{2\theta}}
    = \sqmat{a}{b}{b}{-a}
\]
where $R$ is the magnitude and $\theta$ is the direction, and
$R\geq0 \textrm{ and } \theta\in[0,2\pi)$. Since the matrix is symmetric and
traceless, it is only necessary to store two values, $a$ and $b$ to encode
the tensor.

The major eigenvectors of $t$ are:
\[
    \left\{
        \lambda\mkvec{\cos{\theta}}{\sin{\theta}} ~|~ \lambda \neq 0
    \right\}
\]

The minor eigenvectors are:
\[
    \left\{
        \lambda\mkvec
                {\cos{(\theta+\frac{\pi}{2})}}
                {\sin{(\theta+\frac{\pi}{2})}}
        ~|~ \lambda \neq 0
    \right\}
\]

A tensor field $T$ describes the flow of the constraints across a field, and is
represented as a function that maps from 2D points to tensors. In the context
of this project, a tensor a tool we use to determine the direction of a field
at a given point. This direction is found by calculating the eigenvectors of
the tensors. Major eigenvectors are parallel to the field; minor eigenvectors
are perpendicular.

Another important concept is the hyperstreamline, an effective way to visualize
tensor fields. A hyperstreamline is created by drawing a line through an
eigenvector field such that the line is tangent to every eigenvector along its
path. In this setting, there are major and minor hyperstreamlines that
correspond to the major and minor eigenvectors. The major and minor
eigenvectors of a tensor are guaranteed to be perpendicular for any tensor,
unless it is a degenerate point in the tensor field ($T(\bf p)=0$), for which
eigenvectors cannot be calculated. This property leads to the generation of
perpendicular street intersections, a common feature of real-world road
networks, especially in cities.

\section{Pipeline}\label{sec:pipeline}
We begin with a set of user constraints, which illustrate the general shape of
the roads on the blank canvas. These can be linear or radial, and can be input
as JSON. From these we generate a tensor field, and from that, major and minor
eigenvector fields, which describe the direction stipulated by the constraints
at every point in the area. Finally, based on these eigenvector fields, we
generate seed points from which streamlines are traced across the map. These
streamlines produce the final roads.

\begin{figure*}[t!]
\begin{center}
    \includegraphics[width=5.5in]{images/pipeline.pdf}
\end{center}
\caption{Overview of our pipeline}
\label{fig:pipeline}
\end{figure*}

\subsection{User-Generated Constraints}
We begin with a JSON parser, which accepts linear and radial constraints.
Constraints must be input with their type, linear or radial, in addition to
their position in the domain (in the form of $x$ and $y$ coordinates), and
finally, if the constraint is linear, it’s direction in radians and magnitude.
In addition, the executable will accept width and height parameters
representing the size of the domain to build street maps on. Together, the JSON
constraints and width/height parameters form the initial input data that will
next be used to form tensor fields.

% \floatstyle{plain}
% \restylefloat{figure}
\begin{figure*}[t!]
    \centering
    \vspace{10pt}
    \begin{subfigure}[b]{2.4in}
        \begin{verbatim}
            [{
              "tycon" : "Linear",
              "posx" : 2,
              "posy" : 2,
              "dir" : 0.01,
              "mag" : 3
            },
            {
              "tycon" : "Radial",
              "posx" : 10,
              "posy" : 10,
              "dir" : 0,
              "mag" : 0
            }]
        \end{verbatim}
        \caption{JSON file from by user}
    \end{subfigure}
    \hspace{20pt}
    \begin{subfigure}[b]{2.4in}
        \fbox{\includegraphics[width=\textwidth]{images/json.pdf}}
        \vspace{3pt}
        \caption{generated constraints}
    \end{subfigure}
    \caption{JSON input constraint format}
\label{fig:json}
\end{figure*}

\subsection{Tensor Field Generation}
From each constraint, a tensor field is defined across the entire domain; this
is called a basis field. The basis fields are then combined to produce a tensor
field that describes the space and takes into account all of the specified
constraints. This is done using an exponential decay combinator, such that when
a basis is combined with others, the constraint that defined it has less
influence on the direction of the field the further a point is from it. The
equations in this section follow closely the system detailed in \cite{chen}.

\paragraph{Linear Constraints}
A linear constraint produces the grid pattern that is a common basis for many
cities (consider the streets of Manhattan). The constraint has two parts: a
magnitude $l$ and a direction $\theta$. The basis field for a linear constraint
is defined as:
\[
    T(\pt) =
        l\sqmat{\cos{2\theta}}{\sin{2\theta}}{\sin{2\theta}}{-\cos{2\theta}}
\]
for any point $\bf p$ in the field. It is apparent from this definition that
the direction of a linear basis field is constant, regardless of the value of
$\bf p$.

\paragraph{Radial Constraints}
A radial constraint is used to describe specifically circular areas, such as
roundabouts or more generally curved roads, common in suburban areas. They are
defined at a point $(x_0,y_0)$; major streamlines are drawn as concentric
circles around the point, while minor ones emanate outwards. The basis field
for a radial constraint at a point $\pt=x_p,y_p$ is defined as
\[
    T(\pt) = \sqmat{y^2-x^2}{-2xy}{-2xy}{-(y^2-x^2)}
\]
where $x=x_p-x_0$ and $y=y_p-y_0$.

\paragraph{Combining Basis Fields}
In order to define a tensor field based on multiple basis fields, we use a
function which sums the basis fields together with an exponential falloff,
so the tensor field is most strongly influenced by a basis field close to
the point around which that basis field is defined:
\[
    T(\pt) = \sum_i e^{-d||\pt-\pt_i||^2}T_i(\pt)
\]
where $d$ is a decay constant, $\pt_i$ is the point at which the $i$th
constraint is defined, and $T_i$ is the basis field corresponding to the $i$th
constraint.

\subsection{Eigenvectors}\label{sec:eigenvectors}
The eigenvalues of a tensor $t$ represented by the matrix $\sqmat{a}{b}{b}{-a}$
are:
\begin{align}
    \lambda_1 &= \sqrt{a^2+b^2} \\
    \lambda_2 &= -\lambda_1 \\
\end{align}

The corresponding eigenvectors are:
\[
    \epsilon_i = \left\{1,\frac{\lambda_i-a}{b}\right\}
\]

The eigenvectors are defined if the tensor field is not degenerate (that is,
$a$ and $b$ are both zero) at that point \cite{find-evs}.

\subsection{Streamline Tracing}
Mathematically, the basic streamline tracing algorithm is fairly simple,
however, in order to produce a set of streamlines that approximate a road map,
there are several important cases to consider. In order to trace the
streamlines, we are forced to use a heuristic, since it isn’t possible to trace
a continuous line over a constantly-changing tensor field. In order to
approximate this, we choose a step size $d_\textrm{step}$. Beginning at a seed
point, which is chosen using the methods detailed in Section
\ref{sec:seedpoints}, we calculate the eigenvector at that point, extend a line
in that direction with length $d_\textrm{step}$, and check if we have reached
an end condition.  This process is repeated on the line until one of three end
conditions is reached:

\begin{enumerate}
    \item The point is outside the bounds of the map area
    \item The point is within a certain distance of an existing streamline.
    \item The point is part of a cycle
\end{enumerate}

The first is self-explanatory; if the streamline leaves the boundary of the
map, then we don’t need to draw it any more.

The second is one of the alterations we used, based on the work of \cite{chen}.
When designing a set of roads, it is rare that two streets will approach each
other and not connect, whether they approach at a larger angle (an
intersection) or are running mostly parallel to each other (a merge zone). To
approximate this, at each point we find the nearest neighbor out of the
streamlines that have already been drawn.  If it is within a certain separation
distance, we draw a line directly between the two points.

The third condition arises mainly with radial constraints. Although a true
hyperstreamline would form a perfect circle around a point, our step-size
approximation means that the trace is more likely to form a spiral. This is
obviously a behavior that would not appear in normal maps (other than an
apocalyptic Boston, perhaps). So, when we detect a streamline circling in on
itself, we close the gap and end the spiral.

\begin{figure*}[t!]
    \centering
    \fbox{\includegraphics[width=4in]{images/linear.pdf}}
    \caption{Streamline tracing near road merging}
\end{figure*}

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[b]{2.4in}
        \fbox{\includegraphics[width=\textwidth]{images/cycle_nodetection.pdf}}
        \caption{without cycle detection}
    \end{subfigure}
    \hspace{20pt}
    \begin{subfigure}[b]{2.4in}
        \fbox{\includegraphics[width=\textwidth]{images/cycle_detection.pdf}}
        \caption{with cycle detection}
    \end{subfigure}
\caption{Streamline tracing cycle detection}
\label{fig:cycles}
\end{figure*}

\subsection{Seed Placement}\label{sec:seedpoints}
To begin constructing a streamline, one must choose a seed point—--an initial
coordinate in 2D space from which the streamline is traced. In practice, the
matter of carefully choosing seed points is very important in generating an
evenly spaced configuration of roads. We investigated three methods of placing
seed points:
\begin{enumerate}
    \item Random
    \item Furthest-Point
    \item Even Streamline Spacing ($\chi^2$ distribution)
\end{enumerate}

Purely random placement of seed points generally results in both clusters of
roads which are too close together and large gaps between roads.

In the furthest-point placement, we place each new seed point within the bounds
of the space such that the distance to all existing streamlines is maximized.
As a simplifying heuristic for this method, we generate a number of random
candidate seed points which is small compared to the infinite number of
possible seed points (constrained only by floating point precision) and choose
the one furthest from all existing streamlines. In practice, this offers a
significant improvement over the random seed placement, and results in a more
even distribution of streamlines.

However, even choosing well-spaced seed points will not necessarily result in
an even distribution of streamlines, since two well-spaced seed points can
result in two streamlines which quickly converge. Therefore, we introduce a
third method. To assess how evenly spaced a set of streamlines is, we will
generate a set of discrete points evenly spaced along each streamline, place
these points into $n\times n$ equal-sized buckets covering the bounds of the
space, and perform a chi-squared goodness of fit test on distribution of points
across bucket, where the expected distribution is uniform:
\[
    \chi^2 = \sum_{i=1}^{n^2} \frac{(O_i-E_i)^2}{E_i}
\]
where $n$ is the number of buckets along each axis, $O_i$ is the observed
number of points in the $i$th bucket, and $E_i=\frac{m}{n^2}$ is the
expected number of points in each bucket, assuming a uniform distribution,
where $m$ is the total number of points in the space.

We will choose the seed point whose corresponding streamline, when added to the
existing set of streamlines, optimizes the evenness of streamline placement,
according to the above metric. As with the previous method, we employ a
simplifying heuristic, generating a relatively small set of candidate seed
points, tracing a streamline from each, determining which streamline results in
a more even field of streamlines, and adding it to the existing set of
streamlines.

\section{Assessment}
We set out to achieve two goals: modularity and expressiveness. In this section
we will examine the extent to which these goals were achieved.

\subsection{Modularity and Extensibility}
In Section \ref{sec:seedpoints}, we outlined three methods of seedpoint
placement: random, furthest-point, and $\chi^2$ even spacing. All three
are implemented by functions in the \texttt{Streamline} module, and follow a
similar template. Extending this module with other methods is as simple as
implementing a new function with the same type signature as
\texttt{placeSeeds}, adding a new value constructor to the datatype
\texttt{PlacementMethod}, and adding a new case to the function
\texttt{placeStreamlines} to match the new data constructor. The purity of
functions in a language such as Haskell makes it simple to drop in replacements
for existing functions because one does not have to consider possible side
effects, but only the types of the inputs and output.

In order to detect intersections between streamlines as they are traced, we
defined the \texttt{NearestNeighbor} module to implement a data structure which
stored a set of 2D points and could be queried to determine which of the stored
points was closest to the query point. This is implemented using a type
family, with an instance utilizing a blocked 2D array implementation. This
instance was sufficient for testing purposes, but is currently the performance
bottleneck in streamline tracing. A new instance of the \texttt{Storage}
type family could easily be defined in accordance with the type signature to
provide a more efficient representation. For example, a k-d tree instance of
the type family could offer higher performance. This well-defined abstraction
boundary makes this aspect of our pipeline very easily extensible.

\subsection{Expressiveness}
As our work was not a complete reimplementation of the system described in
\cite{chen}, it is difficult to compare the relative expressiveness of our
Haskell code with the \cpp code of the original implementation. However, we
will present one comparison of functions from our respective implementations
with significantly overlapping functionality. Our Haskell code for calculating
the eigenvectors for a tensor can be found in Figure \ref{fig:evecs}. The
reader may note the close correspondence between the body of these functions
and the mathematical formulas used for calculating eigenvectors in Section
\ref{sec:eigenvectors}.  The 75-line function
\texttt{cal\_eigenvecs\_onevert\_quad} in Chen's implementation. While it is
documented, we found it difficult to determine the side effects it had. In a
language without mutable state or side effect, we found it significantly easier
to fully understand the behavior of functions.  In contrast, even well-written
code in \cpp or another language in the imperative paradigm can be difficult to
understand, because the possibility for side effect requires the reader to have
an understanding of any part of the system that the function can affect, which
in practice can make the effort to understand any small part of a large
codebase much greater.

\begin{figure*}[t!]
\begin{Verbatim}[frame=single]
-- returns a tuple of the (larger, smaller) eigenvectors of a tensor
eigenvalues :: Tensor -> (Float, Float)
eigenvalues (Tensor a b d) =
let eval = sqrt (a*a + b*b)
in  (eval, (-1)*eval)

-- returns a tuple of the (major, minor) eigenvectors for a tensor
eigenvectors :: Tensor -> (V2.Vector2, V2.Vector2)
eigenvectors (Tensor a 0.0 d) =
if a > d then (V2.Vector2 1 0, V2.Vector2 0 1)
   else (V2.Vector2 0 1, V2.Vector2 1 0)
   eigenvectors (Tensor a b d) =
     let (eval1, eval2) = eigenvalues (Tensor a b d)
         mkEvec eval    = V2.Vector2 1 ((eval - a) / b)
     in (mkEvec eval1, mkEvec eval2)
\end{Verbatim}

\caption{Haskell implementation of eigenvector calculation (from
            \texttt{Tensor.hs})}
\label{fig:evecs}
\end{figure*}

\section{Conclusion}
In this paper we have presented the underlying mathematics and functional
implementation of the work described in \cite{chen}, with the goals being that
a functional implementation would readily express the mathematics, provide
modularity that the imperative implementation provide in \cite{chen} could not.
We believe that, due to the focus on expressibility and modularity, assessed in
the previous section, that an individual looking to understand the pipeline
will find our code undestandable and approachable, and lend itself to be
easily extended. We hope that our work will motivate a greater use of
functional languages in future efforts in the research of new procedural
algorithms.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}